---
title: "CRM Migration - TMC Interview"
author: "Joe Ciesielski"
date: "2019-05-28"
output:
  ioslides_presentation:
    css: www/style.css
    incremental_bullets: true
---

```{r setup, include = FALSE}

# Your member has decided to switch CRMs (the system where they organize their email list, events, and other volunteer information). They have asked for TMC’s support to complete this work. Please outline:
# What questions might you ask the member at the outset of this migration?
# How would you organize yourself to project manage the migration?
# What kinds of things could go poorly?
# We want to better understand your approach to this hypothetical task and we will discuss during our call. You will have 5 minutes to outline your process, feel free to come with any tools (Asana, Trello, etc), documentation (1 pager), spreadsheet, Powerpoint or anything you may need to support the conversation. You will be able to share your screen.
```

```{r, echo=FALSE, eval=FALSE}

rmarkdown::render(here::here("crm.Rmd"))

```
## Key questions

* Principles
  * Core values? Organizational mission? 
* Purpose
  * What is the purpose of CRM? Why change? 
* Vision and outcomes
  * What are you hoping to achieve with new tool? 
* Immediate actions
  * What has your attention? 
<br>
<br>
* Work from the bottom up (*GTD* approach)

## Project Management

* Process driven by goals and needs
    * find the right balance of lean and waterfall
* Tools integrated into workflow flow from approach
    * options: Trello, Gantt, Basecamp
* Identify key staff and training needs
* Establish project brief, timeline, and communication 

## Potential Problems

* Premature decision-making / optimization
* Lack of clarity around key decision points
    * e.g. mapping fields inappropriately
* Connections to existing systems
* Customizations of exisiting or new tools
